Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-03-11 cis-perfect-testing.tex
fix subscripts for survival analysis

% \section{Definitions}
% \todo[inline]{These will be defined in the introductory chapters, but are here so that this draft is understandable}
% \begin{itemize}
% \item
%   We assume that, for all episodes $i$, $D_i$ is iid and independent
%   of the time of the infection. Define the survival function
%   $\prob(D_i \geq t \mid B_i = b, \theta) = \prob(D_i \geq t \mid \theta) = S_\theta(t)$,
%   where $\theta$ are the parameters controlling the survival
%   distribution (the discussion in this document is valid regardless of
%   the model specified for $S_\theta$ and hence we consider $\theta$
%   as an arbitrary vector of parameters).
% \item
%   The beginning of episode $i$ is known to occur in the interval
%   $[l_i^{(b)}, r_i^{(b)}]$, and similarly for the end of the infection
%   in $[l_i^{(e)}, r_i^{(e)}]$.
% \end{itemize}

@@

The second complication is that the exact beginning and end times of detected infection episodes are unknown.
The imprecision in the beginning of the infection episode arises because the only observation that an episode has started is that an individual was negative at one test, on day $l_i^{(b)}-1$ (the minus one is because the first day that the infection could have begun is the day after the negative test), and positive at the next test, on day $r_i^{(b)}$.
Similarly, the end of the infection episode is bounded by the final positive test on day $l_i^{(e)}$ and the following negative test on day $r_i^{(e)}+1$ (the plus one is because the last day that the infection could have ended is the day before the negative test).
\Cref{perf-test:fig:double-interval-censor} shows this issue graphically.
\begin{figure}
  \centering \includegraphics{cis-perfect-testing/double-interval-censor}
  \caption[Double-interval censoring in CIS data]{Episodes data in CIS is double interval censored, meaning that both the start and end of the episode are only known up to an interval. Demonstrated here by a participant who is recorded negative at time 7 to positive at time 14, bounding the start of the episode; similarly, changing from positive at time 28 to negative at time 56 bounds the end of the episode. \label{perf-test:fig:double-interval-censor}}
\end{figure}

@@

Integrating over the number of missed episodes is possible analytically.
This makes inference tractable within modern, high-performance Bayesian software such as Stan.
Let the total number of infection episodes in the cohort be $\ntot = \ndet + \nnodet$ where $\ndet$ is the number of detected infections and $\nnodet$ is the number of undetected infections.
Let $\Ncis$ be the number of individuals in the cohort and $n_\sched$ be the number of test schedules.
%Without loss of generality, let the individuals with a detected episode be indexed by $i = 1, \dots, \ndet$ and the individuals without a detected episode be indexed by $i = \ndet + 1, \dots, \Ncis$.
Let $i = 1, \dots \ntot$ index infection episodes.

All the relevant information about episode $i$ can be fully characterised by the (unobserved) triplet $(b_i, e_i, \sched_i)$ where $b_i$ and $e_i$ are the times the episode being and end respectively and $\sched_i$ is the set of test times for the individual in which the episode occurs.
The episode's duration, the number of days the individual in which the episode occurred, was positive for is $d_i = e_i - b_i + 1$.
This triplet is assumed iid.
If $\sched_i$ contains at least one time between $b_i$ and $e_i$ (inclusive), one time prior to $b_i$, and one time after $e_i$, then the episode is detected, otherwise it is undetected.
$b_i$ and $e_i$ span possible infection and recovery times.
$\sched_i$ is one of the observed test schedules, with any other test schedule assumed to occur with probability zero.
Over the short period of time considered, each unique test schedule has zero or one detected episodes.
Let $\set{D}$ be the set of test schedules with a detected infection in the cohort.

To relate the state space to the data, I define three subsets partitioning the space of possible infection and recovery times (\ie: are strict subsets of $T \times T$), for each test schedule.
Figure \ref{perf-test:fig:partitionSpace} shows this graphically.
Each episode, whether detected or not, must fall into one of these three classes with a probability that is a function of the model parameters, $\theta$.

\begin{itemize}
\item
  Admissible episodes, $\alpha_i$, which have an infection and recovery time consistent with the data observed.
  This space is empty for $i \notin \set{D}$.
  If $i \in \set{D}$ then $n_{ia} =1$ admissible episodes occurred in an individual with test schedule $\sched_{i}$, otherwise $n_{ia} = 0$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, it is admissible with probability $p_{ia} = \prob((b, e) \in \alpha_i \mid \theta)$.
  \label{perf-test:def:admissible}
\item
  Undetected episodes, $\Omega_i^C$, which have an infection and recovery time such that they would not have tested positive.
  An unknown number, $n_{iu}$ of these observed in individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, it is undetected with probability $p_{iu} = \prob((b, e) \in \Omega^C_i \mid \theta)$.
\item
  Inadmissible episodes, $\beta_i$, which are not consistent with the data observed but would have been
  detected.
  This is all remaining episodes not in the previous sets.
  $n_{i\bar{a}} = 0$ of these occurred in any individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in these individuals, it is inadmissible with probability $p_{i\bar{a}} = \prob((b, e) \in \beta_i \mid \theta)$.
\end{itemize}

@@

These three classes span all possible episodes and are mutually exclusive.
Hence, $\ntot = \ndet + \nnodet + n_{i\bar{a}} = \sum_{i=1}^{\ndet} n_{ia} + \sum_{i=1}^{\ntot} n_{iu} + 0$.
Furthermore, for all $i$, $p_{ia} + p_{iu} + p_{i\bar{a}} = 1$.

Any episode $i$ independently belongs to one of these classes.
I assume the individual in which each episode occurs is iid uniform across the cohort.
Therefore, conditional on $\ntot$, the number of episodes in each class, $(n_{ia}, n_{iu}, n_{i\bar{a}})^T$, are distributed multinomially.
The above probabilities ($p_{iu}$, $p_{ia}$, and $p_{i\bar{a}}$) are conditional on the episode occurring with test schedule $\sched_i$.
These probabilities are multiplied by $\eta_i/\Ncis$, where $\eta_i$ is the number of individuals with test schedule $\sched_i$, to give the probability that an episode belongs to each class, without conditioning on a specific individual.

The undetected classes for each individual cannot be distinguished by any observable data.
Each class's count needs to be marginalised over to calculate the likelihood.
Because the counts have a joint multinomial distribution, this is equivalent to only considering the sum of the counts.
The overall probability of an episode being unobserved is $\pnodet = \frac{1}{\Ncis} \sum_{i=1}^{n_\sched} \eta_i p_{iu}$.
Furthermore, the inadmissible classes are guaranteed to have zero episodes.
Counts of zero do not contribute to a multinomial pmf (they multiply by 1).
Therefore, I combine the undetected classes into a single class and ignore the inadmissible classes.

Let $\na = (n_{1a}, \dots, n_{n_\sched a})^T$ be the vector of counts of episodes in the admissible classes.
The counts $\na$ and $\nnodet$ are jointly distributed multinomially.
The support of the distribution requires $\nnodet = \ntot - \sum_{i=1}^{n_\sched} n_{ia} = \ntot - \ndet$ and $\ntot \geq \ndet$.
Hence:
\begin{align}
&p(\na \mid \ntot, \theta) \\
&= p(\nnodet, \na\mid \ntot, \theta) \\
&= \frac{\ntot!}{\nnodet! \prod_{i=1}^{n_\sched} n_{ia}!}  \pnodet^{\nnodet}\prod_{i=1}^{n_\sched} \left(\frac{1}{n_\sched}p_{ia}\right)^{n_{ia}} \\
&\propto \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \label{perf-test:eq:augmented-likelihood}
\end{align}
the final line follows because $\nnodet = \ntot - \ndet$, $n_{ia} = 1$ for $i \in \set{D}$, and $n_{ia} = 0$ otherwise.

@@

The posterior density we are interested in is $p(\theta \mid \na)$.
It can be written as follows.
\begin{align}
&p(\theta \mid \na) \\
&\propto p(\theta) p(\na \mid \theta) \\
&= p(\theta) \sum_{\ntot=\ndet}^\infty p(\ntot \mid \theta) p(\na \mid \theta, \ntot) \\
&\propto p(\theta) \sum_{\ntot=\ndet}^\infty \left( p(\ntot \mid \theta) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i=\ndet}^{\ndet} p_{ia} \right) \\
&= p(\theta) \left( \prod_{i=\ndet}^{\ndet} p_{ia} \right) \left( \sum_{\ntot=\ndet}^\infty p(\ntot \mid \theta) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \right) \label{perf-test:eq:full-posterior}.
\end{align}
The rest of this section derives expressions for each $p_{ia}$, $p_{u}$ and the sum.

@@

Substituting this into \cref{perf-test:eq:full-posterior} gives:
\begin{align}
p(\theta \mid \na)
&\propto p(\theta) \left( \prod_{i=\ndet}^{\ndet} p_{ia} \right) (r + \mu (1- \pnodet))^{-(r+\ndet)} \label{perf-test:eq:full-posterior-simplified}.
\end{align}

@@

Next, I derive $p_{ia}$.
By definition (see~\cpageref{perf-test:def:admissible}), we have:
\begin{align}
p_{ia} &= \prob((b, e) \in \alpha_i \mid \theta) \\
\alpha_i &= \{ (b, e) : l_i^{(b)} \leq b \leq r_i^{(b)} \wedge l_i^{(e)} \leq e \leq r_i^{(e)}\}.
\intertext{Suppressing the conditioning on $\theta$, this gives:}
p_{ia}
&= \prob \left( l_i^{(b)} \leq B_i \leq r_i^{(b)}, l_i^{(e)} \leq E_i \leq r_i^{(e)} \right) \\
&= \prob \left( l_i^{(e)} \leq E_i \leq r_i^{(e)} \mid l_i^{(b)} \leq B_i \leq r_i^{(b)} \right) \prob \left( l_i^{(b)} \leq B_i \leq r_i^{(b)} \right) \\
&=\sum_{b = l_i^{(l)}}^{r_i^{(b)}} \prob \left( l_i^{(e)} \leq E_i \leq r_i^{(e)} \mid B_i = b \right) \prob \left(B_i = b \right) \\
&=\sum_{b = l_i^{(b)}}^{r_i^{(b)}} \prob \left( l_i^{(e)} - b + 1 \leq D_i \leq r_i^{(e)} - b + 1 \right) \prob \left(B_i = b \right) &\text{by def of $D_i$} \\
&=\sum_{b = l_i^{(b)}}^{r_i^{(b)}} \left( S_\theta(l_i^{(e)} - b + 1) - S_\theta(r_i^{(e)} - b + 1) \right) \prob \left(B_i = b \right) &\text{by def of $S_\theta$} \\
&\propto \sum_{b = l_i^{(b)}}^{r_i^{(b)}} \left( S_\theta(l_i^{(e)} - b + 1) - S_\theta(r_i^{(e)} - b + 1) \right)
\end{align}
under the assumption of uniform probability of infection time.

The next step is to derive $1 - p_u = 1 - \frac{1}{N} \sum_{i=1}^{\Ncis} p_{iu} = \frac{1}{N} \sum_{i=1}^{\Ncis} (1 - p_{iu})$.
Therefore, the crucial component is $1 - p_{iu} = 1 - \prob((B_i, E_i) \in \Omega^C_i \mid \theta)$, one minus the probability that episode $i$ was undetected.
An episode is undetected if and only if no tests are performed during the episode or if there was no negative test prior to the episode.
Equivalently, that the first test at or after $b$ is after $e$, or that there is no negative test prior to $b$.
To formalise this, first define $\tau_{\sched_i}(t)$ as the time until the next test at or after time $t$ in the schedule $\sched_i$:
\begin{align}
\tau_{\sched_i}(t) &= \min \{ t' \in \sched_i : t' \geq t \} - t.
\end{align}
Then $\Omega^C_i$ can be written as:
\begin{align}
\Omega_i^C = \{ (b, e) \ssep \tau_{\sched_i}(b) + b > e \vee b \leq \min(\sched_i) \}.
\end{align}
Therefore, suppressing the conditioning on $\theta$:
\begin{align}
1 - p_{iu}
&= 1 - \prob((B_i, E_i) \in \Omega^C_i) \\
&= 1 - \prob( \tau_{\sched_i}(B_i) + B_i > E_i \vee B_i \leq \min(\sched_i) ) \\
&= 1 - \prob( \tau_{\sched_i}(B_i) + 1 > D_i , B_i > \min(\sched_i) )  - \prob( B_i \leq \min(\sched_i) ) \\
&= 1 - \sum_{b={\min\sched_i} + 1}^T \prob( \tau_{\sched_i}(b) + 1> D_i) \prob(B_i = b) - \sum_{t=1}^{\min\sched_i} \prob(B_i = b)\\
&= 1 - \frac{1}{T} \sum_{b=\min(\sched_i)+1}^T (1 - S_\theta(\tau_{\sched_i}(b) + 1)) - \frac{1}{T} \min\sched_i \\
&= \frac{1}{T} \left(T - \sum_{b=\min\sched_i+1}^T (1 - S_\theta(\tau_{\sched_i}(b) + 1)) - \min\sched_i \right) \\
&= \frac{1}{T} \left(T - (T - \min\sched_i) + \sum_{b=\min\sched_i+1}^T S_\theta(\tau_{\sched_i}(b + 1)) - \min\sched_i \right) \\
&= \frac{1}{T} \sum_{b=\min(\sched_i)+1}^T S_\theta(\tau_{\sched_i}(b) + 1). \label{perf-test:eq:piu}
\end{align}

@@

I simulate a dataset of detected episodes that has the same characteristics as that in the CIS by the following procedure.
\begin{enumerate}
    \item Extract the test schedules for each individual who had at least one test during the period of interest.
    \item Draw an episode start time, $b_i$ for each individual uniformly at random between 2nd July 2020 (100 days before the period where a detected episode would be included) and 6th December 2020 (the end of this period).
    \item Draw a duration of episode for each individual, $d_i$, based on the distribution described below. Then calculate the end of their infection episode, $e_i = b_i + d_i - 1$.
    \item Simulate the test results based on the test schedule, $b_i$, and $e_i$. Tests between $b_i$ and $e_i$ (inclusive) are positive, all other tests are negative.
    \item Discard episodes where there are no positive tests (\ie undetected episodes) and then apply the same inclusions/exclusions as in \cref{perf-test:sec:problem}.
    \item Of these remaining episodes, sample 4,800 to match the sample size of the true dataset. This is needed because in step 2 the entire cohort was infected, while in the real study only a (unknown) portion is infected.
    \item For this final set of episodes, calculate $(l_i^{(b)}, r_i^{b}, l_i^{(e)}, r_i^{(e)})$ by taking the day after the last negative prior to any positives, the first positive, the last positive, and the day before the negative following the last positive respectively.
\end{enumerate}

>>>>

% \section{Definitions}
% \todo[inline]{These will be defined in the introductory chapters, but are here so that this draft is understandable}
% \begin{itemize}
% \item
%   We assume that, for all episodes $i$, $D_i$ is iid and independent
%   of the time of the infection. Define the survival function
%   $\prob(D_i \geq t \mid B_{j(i)} = b, \theta) = \prob(D_i \geq t \mid \theta) = S_\theta(t)$,
%   where $\theta$ are the parameters controlling the survival
%   distribution (the discussion in this document is valid regardless of
%   the model specified for $S_\theta$ and hence we consider $\theta$
%   as an arbitrary vector of parameters).
% \item
%   The beginning of episode $i$ is known to occur in the interval
%   $[l_{j(i)}^{(b)}, r_{j(i)}^{(b)}]$, and similarly for the end of the infection
%   in $[l_{j(i)}^{(e)}, r_{j(i)}^{(e)}]$.
% \end{itemize}

@@

The second complication is that the exact beginning and end times of detected infection episodes are unknown.
The imprecision in the beginning of the infection episode arises because the only observation that episode $j$ has started is that an individual was negative at one test, on day $l_j^{(b)}-1$ (the minus one is because the first day that the infection could have begun is the day after the negative test), and positive at the next test, on day $r_j^{(b)}$.
Similarly, the end of the infection episode is bounded by the final positive test on day $l_j^{(e)}$ and the following negative test on day $r_j^{(e)}+1$ (the plus one is because the last day that the infection could have ended is the day before the negative test).
\Cref{perf-test:fig:double-interval-censor} shows this issue graphically.
\begin{figure}
  \centering \includegraphics{cis-perfect-testing/double-interval-censor}
  \caption[Double-interval censoring in CIS data]{Episodes data in CIS is double interval censored, meaning that both the start and end of the episode are only known up to an interval. Demonstrated here by a participant who is recorded negative at time 7 to positive at time 14, bounding the start of the episode; similarly, changing from positive at time 28 to negative at time 56 bounds the end of the episode. \label{perf-test:fig:double-interval-censor}}
\end{figure}

@@

Integrating over the number of missed episodes is possible analytically.
This makes inference tractable within modern, high-performance Bayesian software such as Stan.
Let the total number of infection episodes in the cohort be $\ntot = \ndet + \nnodet$ where $\ndet$ is the number of detected infections and $\nnodet$ is the number of undetected infections.
Let $\Ncis$ be the number of individuals in the cohort and $\nsched$ be the number of test schedules.
%Without loss of generality, let the individuals with a detected episode be indexed by $i = 1, \dots, \ndet$ and the individuals without a detected episode be indexed by $i = \ndet + 1, \dots, \Ncis$.
Let $i = 1, \dots, \nsched$ index infection schedules and $j = 1, \dots \ntot$ index infection episodes.

All the relevant information about episode $i$ can be fully characterised by the (unobserved) triplet $(b_j, e_j, \sched_j)$ where $b_j$ and $e_j$ are the times the episode being and end respectively and $\sched_j$ is the set of test times for the individual in which the episode occurs.
The episode's duration is $d_j = e_j - b_j + 1$.
%, the number of days the individual in which the episode occurred, was positive for 
This triplet is assumed iid.

$b_j$ and $e_j$ span possible infection and recovery times.
$\sched_j$ is one of the observed test schedules, with any other test schedule assumed to occur with probability zero.
Over the short period of time considered, each unique test schedule has zero or one detected episodes.

Let $i \in \set{D}$ indicate $\sched_i$ is associated with a detected episode.
That is, $i \in \set{D}$ if and only if there exists a $j(i)$ such that $b_{j(i)}$ and $e_{j(i)}$ such that $(b_{j(i)}, e_{j(i)}, \sched_i)$ is a detected episode.
$(b_{j(i)}, e_{j(i)}, \sched_i)$ is a detected episode when the following three conditions are met.
First, $\exists t \in \sched_i$ such that $b_{j(i)} \leq t \leq e_{j(i)}$; this condition is equivalent to having at least one positive test for the episode.
Second, $\exists t \in \sched_i$ such that $t < b_{j(i)}$, to lower bound the start of the episode.
Finally, $\exists t \in \sched_i$ such that $t > e_{j(i)}$, to upper bound the end of the episode; this condition will be met for any episode sufficiently far in the past in individuals not lost to follow-up (almost always the case in this context).
For a new context where recent infections need to be considered, this condition can be relaxed by considering these episodes as right-censored, at little cost except a more complicated likelihood expression.

To relate the state space to the data, I define three subsets partitioning the space of possible infection and recovery times (\ie: are strict subsets of $T \times T$), for a test schedule $\sched_i$.
Figure \ref{perf-test:fig:partitionSpace} shows this graphically.
Each episode, whether detected or not, must fall into one of these three classes with a probability that is a function of the model parameters, $\theta$.

\begin{itemize}
\item
  Admissible episodes, $\alpha_i$, which have an infection and recovery time consistent with the data observed.
  This space is empty for $i \notin \set{D}$; in this case, $n_{ia} = 0$ admissible episodes occurred in individuals with test schedule $\sched_i$.
  Otherwise, $i \in \set{D}$ then $n_{ia} =1$ admissible episode occurred in an individual with test schedule $\sched_{i}$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, it is admissible with probability $p_{ia} = \prob((b, e) \in \alpha_i \mid \theta)$.
  \label{perf-test:def:admissible}
\item
  Undetected episodes, $\Omega_i^C$, which have an infection and recovery time such that they would not have tested positive.
  An unknown number, $n_{iu}$ of these observed in individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, it is undetected with probability $p_{iu} = \prob((b, e) \in \Omega^C_i \mid \theta)$.
\item
  Inadmissible episodes, $\beta_i$, which are not consistent with the data observed but would have been
  detected.
  This is all remaining episodes not in the previous sets.
  $n_{i\bar{a}} = 0$ of these occurred in any individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in these individuals, it is inadmissible with probability $p_{i\bar{a}} = \prob((b, e) \in \beta_i \mid \theta)$.
\end{itemize}

@@

These three classes span all possible episodes and are mutually exclusive.
Hence, $\ntot = \ndet + \nnodet + n_{i\bar{a}} = \sum_{i=1}^{\nsched} n_{ia} + \sum_{i=1}^{\nsched} n_{iu} + 0$.
Furthermore, for all $i$, $p_{ia} + p_{iu} + p_{i\bar{a}} = 1$.

Any episode $j$ independently belongs to one of these classes.
I assume the individual in which each episode occurs is iid uniform across the cohort.
Therefore, conditional on $\ntot$, the number of episodes in each class, $(n_{ia}, n_{iu}, n_{i\bar{a}})^T$, are distributed multinomially.
The above probabilities ($p_{iu}$, $p_{ia}$, and $p_{i\bar{a}}$) are conditional on the episode occurring with test schedule $\sched_i$.
These probabilities are multiplied by $\eta_i/\Ncis$, where $\eta_i$ is the number of individuals with test schedule $\sched_i$, to give the probability that an episode belongs to each class, without conditioning on a specific individual.

The undetected classes for each individual cannot be distinguished by any observable data.
Each class's count needs to be marginalised over to calculate the likelihood.
Because the counts have a joint multinomial distribution, this is equivalent to only considering the sum of the counts.
The overall probability of an episode being unobserved is $\pnodet = \frac{1}{\Ncis} \sum_{i=1}^{\nsched} \eta_i p_{iu}$.
Furthermore, the inadmissible classes are guaranteed to have zero episodes.
Counts of zero do not contribute to a multinomial pmf (they multiply by 1).
Therefore, I combine the undetected classes into a single class and ignore the inadmissible classes.

Let $\na = (n_{1a}, \dots, n_{\nsched a})^T$ be the vector of counts of episodes in the admissible classes.
The counts $\na$ and $\nnodet$ are jointly distributed multinomially.
The support of the distribution requires $\nnodet = \ntot - \sum_{i=1}^{\nsched} n_{ia} = \ntot - \ndet$ and $\ntot \geq \ndet$.
Hence:
\begin{align}
&p(\na \mid \ntot, \theta) \\
&= p(\nnodet, \na\mid \ntot, \theta) \\
&= \frac{\ntot!}{\nnodet! \prod_{i=1}^{\nsched} n_{ia}!}  \pnodet^{\nnodet}\prod_{i=1}^{\nsched} \left(\frac{1}{\nsched}p_{ia}\right)^{n_{ia}} \\
&\propto \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \label{perf-test:eq:augmented-likelihood}
\end{align}
the final line follows because $\nnodet = \ntot - \ndet$, $n_{ia} = 1$ for $i \in \set{D}$, and $n_{ia} = 0$ otherwise.

@@

The posterior density we are interested in is $p(\theta \mid \na)$.
It can be written as follows.
\begin{align}
&p(\theta \mid \na) \\
&\propto p(\theta) p(\na \mid \theta) \\
&= p(\theta) \sum_{\ntot=\ndet}^\infty p(\ntot \mid \theta) p(\na \mid \theta, \ntot) \\
&\propto p(\theta) \sum_{\ntot=\ndet}^\infty \left( p(\ntot \mid \theta) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \right) \\
&= p(\theta) \left( \prod_{i \in \set{D}} p_{ia} \right) \left( \sum_{\ntot=\ndet}^\infty p(\ntot \mid \theta) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \right) \label{perf-test:eq:full-posterior}.
\end{align}
The rest of this section derives expressions for each $p_{ia}$, $p_{u}$ and the sum.

@@

Substituting this into \cref{perf-test:eq:full-posterior} gives:
\begin{align}
p(\theta \mid \na)
&\propto p(\theta) \left( \prod_{i \in \set{D}} p_{ia} \right) (r + \mu (1- \pnodet))^{-(r+\ndet)} \label{perf-test:eq:full-posterior-simplified}.
\end{align}

@@

Next, I derive $p_{ia}$.
By definition (see~\cpageref{perf-test:def:admissible}), we have, for any $i \in \set{D}$:
\begin{align}
p_{ia} &= \prob((b, e) \in \alpha_i \mid \theta) \\
\alpha_i &= \{ (b, e) : l_{j(i)}^{(b)} \leq b \leq r_{j(i)}^{(b)} \wedge l_{j(i)}^{(e)} \leq e \leq r_{j(i)}^{(e)}\}.
\intertext{Suppressing the conditioning on $\theta$, this gives:}
p_{ia}
&= \prob \left( l_{j(i)}^{(b)} \leq B_{j(i)} \leq r_{j(i)}^{(b)}, l_{j(i)}^{(e)} \leq E_{j(i)} \leq r_{j(i)}^{(e)} \right) \\
&= \prob \left( l_{j(i)}^{(e)} \leq E_{j(i)} \leq r_{j(i)}^{(e)} \mid l_{j(i)}^{(b)} \leq B_{j(i)} \leq r_{j(i)}^{(b)} \right) \prob \left( l_{j(i)}^{(b)} \leq B_{j(i)} \leq r_{j(i)}^{(b)} \right) \\
&=\sum_{b = l_{j(i)}^{(l)}}^{r_{j(i)}^{(b)}} \prob \left( l_{j(i)}^{(e)} \leq E_{j(i)} \leq r_{j(i)}^{(e)} \mid B_{j(i)} = b \right) \prob \left(B_{j(i)} = b \right) \\
&=\sum_{b = l_{j(i)}^{(b)}}^{r_{j(i)}^{(b)}} \prob \left( l_{j(i)}^{(e)} - b + 1 \leq D_{j(i)} \leq r_{j(i)}^{(e)} - b + 1 \right) \prob \left(B_{j(i)} = b \right) &\text{by def of $D_{j(i)}$} \\
&=\sum_{b = l_{j(i)}^{(b)}}^{r_{j(i)}^{(b)}} \left( S_\theta(l_{j(i)}^{(e)} - b + 1) - S_\theta(r_{j(i)}^{(e)} - b + 1) \right) \prob \left(B_{j(i)} = b \right) &\text{by def of $S_\theta$} \\
&\propto \sum_{b = l_{j(i)}^{(b)}}^{r_{j(i)}^{(b)}} \left( S_\theta(l_{j(i)}^{(e)} - b + 1) - S_\theta(r_{j(i)}^{(e)} - b + 1) \right)
\end{align}
under the assumption of uniform probability of infection time.

The next step is to derive $1 - p_u = 1 - \frac{1}{N} \sum_{i=1}^{\nsched} p_{iu} = \frac{1}{N} \sum_{i=1}^{\nsched} (1 - p_{iu})$.
Therefore, the crucial component is $1 - p_{iu} = 1 - \prob((B_{j(i)}, E_{j(i)}) \in \Omega^C_i \mid \theta)$, one minus the probability that episode $i$ was undetected.
An episode is undetected if and only if no tests are performed during the episode or if there was no negative test prior to the episode.
Equivalently, that the first test at or after $b$ is after $e$, or that there is no negative test prior to $b$.
To formalise this, first define $\tau_{\sched_i}(t)$ as the time until the next test at or after time $t$ in the schedule $\sched_i$:
\begin{align}
\tau_{\sched_i}(t) &= \min \{ t' \in \sched_i : t' \geq t \} - t.
\end{align}
Then $\Omega^C_i$ can be written as:
\begin{align}
\Omega_i^C = \{ (b, e) \ssep \tau_{\sched_i}(b) + b > e \vee b \leq \min(\sched_i) \}.
\end{align}
Therefore, suppressing the conditioning on $\theta$:
\begin{align}
1 - p_{iu}
&= 1 - \prob((B_{j(i)}, E_{j(i)}) \in \Omega^C_i) \\
&= 1 - \prob( \tau_{\sched_i}(B_{j(i)}) + B_{j(i)} > E_{j(i)} \vee B_{j(i)} \leq \min(\sched_i) ) \\
&= 1 - \prob( \tau_{\sched_i}(B_{j(i)}) + 1 > D_i , B_{j(i)} > \min(\sched_i) )  - \prob( B_{j(i)} \leq \min(\sched_i) ) \\
&= 1 - \sum_{b={\min\sched_i} + 1}^T \prob( \tau_{\sched_i}(b) + 1> D_i) \prob(B_{j(i)} = b) - \sum_{t=1}^{\min\sched_i} \prob(B_{j(i)} = b)\\
&= 1 - \frac{1}{T} \sum_{b=\min(\sched_i)+1}^T (1 - S_\theta(\tau_{\sched_i}(b) + 1)) - \frac{1}{T} \min\sched_i \\
&= \frac{1}{T} \left(T - \sum_{b=\min\sched_i+1}^T (1 - S_\theta(\tau_{\sched_i}(b) + 1)) - \min\sched_i \right) \\
&= \frac{1}{T} \left(T - (T - \min\sched_i) + \sum_{b=\min\sched_i+1}^T S_\theta(\tau_{\sched_i}(b + 1)) - \min\sched_i \right) \\
&= \frac{1}{T} \sum_{b=\min(\sched_i)+1}^T S_\theta(\tau_{\sched_i}(b) + 1). \label{perf-test:eq:piu}
\end{align}

@@

I simulate a dataset of detected episodes that has the same characteristics as that in the CIS by the following procedure.
\begin{enumerate}
    \item Extract the test schedules for each individual who had at least one test during the period of interest.
    \item Draw an episode start time, $b_{j(i)}$ for each individual uniformly at random between 2nd July 2020 (100 days before the period where a detected episode would be included) and 6th December 2020 (the end of this period).
    \item Draw a duration of episode for each individual, $d_i$, based on the distribution described below. Then calculate the end of their infection episode, $e_{j(i)} = b_{j(i)} + d_i - 1$.
    \item Simulate the test results based on the test schedule, $b_{j(i)}$, and $e_{j(i)}$. Tests between $b_{j(i)}$ and $e_{j(i)}$ (inclusive) are positive, all other tests are negative.
    \item Discard episodes where there are no positive tests (\ie undetected episodes) and then apply the same inclusions/exclusions as in \cref{perf-test:sec:problem}.
    \item Of these remaining episodes, sample 4,800 to match the sample size of the true dataset. This is needed because in step 2 the entire cohort was infected, while in the real study only a (unknown) portion is infected.
    \item For this final set of episodes, calculate $(l_j^{(b)}, r_j^{b}, l_j^{(e)}, r_j^{(e)})$ by taking the day after the last negative prior to any positives, the first positive, the last positive, and the day before the negative following the last positive respectively.
\end{enumerate}

### 2024-03-11 symbols.tex
fix subscripts for survival analysis

From chapter 6
\begin{itemize}
    \item $p_{ia}'$ the replacement for $p_{ia}$ including false negatives
    \item $p_{iu}'$ the replacement for $p_{iu}$ including false negatives
    \item $\psens$ test sensitivity
    \item $r\inform$ the dispersion of the informative prior on $\ntot$
    \item $y_i(t)$ the test result for infection episode $i$ at time $t$
    \item $\vec{y_i}$ the test results associated with infection episode $i$
    \item $\eta$ the additional probability of missing an infection episode due to false negatives
    \item $\mu\inform$ the mean of the informative prior on $\ntot$
    \item $\tau_{\sched_i}(t)$ the time between $t$ and the next but one test for test schedule $i$
\end{itemize}

>>>>

From chapter 6
\begin{itemize}
    \item $p_{ia}'$ the replacement for $p_{ia}$ including false negatives
    \item $p_{iu}'$ the replacement for $p_{iu}$ including false negatives
    \item $\psens$ test sensitivity
    \item $r\inform$ the dispersion of the informative prior on $\ntot$
    \item $y_{j(i)}(t)$ the test result for infection episode $i$ at time $t$
    \item $\vec{y_{j(i)}}$ the test results associated with infection episode $j(i)$ with test schedule $\sched_i$
    \item $\eta$ the additional probability of missing an infection episode due to false negatives
    \item $\mu\inform$ the mean of the informative prior on $\ntot$
    \item $\tau_{\sched_i}(t)$ the time between $t$ and the next but one test for test schedule $i$
\end{itemize}
